Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-11-17 Simulation.tex
edits for figures and equations finished, working on scope and noise

After the computation of the short time Fourier transform; the Doppler envelope is calculated by performing a threshold on the power, in both the positive and negative Doppler, at each time slice. The envelope value for that time slice is found by approaching zero frequency and stopping as the power reaches a predetermined level from both the negative and positive values. The power level to set the threshold is calculated differently based on whether the elevation or the azimuth parameter is being swept. For the azimuth angle sweep, which will be shown later in section \ref{taa}, the maximum power is found in each time slice then those values are averaged. Then for each successive angle the same average is taken and averaged with all the others.

@@

The Doppler envelopes are displayed in Figure \ref{fig:test_spec_w_doppler_profile}, the lower in blue and the upper in red. The minimum is calculated from the lower envelope, and the maximum from the upper envelope. Those values are then used to characterize that Doppler profile and analyze signal trends as different variables change.

@@

The rotor blade length from Figure \ref{fig:rx_position_image} is 7.5 m so the receiver will be varied from 0 m to 9 m in 1 m increments. The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of the position of the receiver.

%figure of maximum doppler vs rx position
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/receiver_position_max_doppler.eps}
		\caption{Max and Min Doppler vs. Receiver Position with Tx Azimuth at 135\textdegree \space and Elevation at 54\textdegree}
		\label{fig:rx_position}
	\end{center}
\end{figure}
%azimuth = 135, elevation = 54deg

Figure \ref{fig:rx_position} shows maximum Doppler versus the position of the receiver. From this graph, the maximum Doppler frequency changes significantly as the receiver is moved from directly underneath the rotor to past the tip of the rotor at 9 m.The position of the receiver also effects the resulting signal envelope shapes shown in Figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

%figure of spectrogram at 0m -- figure of spectrogram at 7m side by side
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Doppler_Receiver_0m.eps}
		\caption{Doppler Envelope with receiver at 0m}
		\label{fig:rx_position_0m}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Doppler_Receiver_7m.eps}
		\caption{Doppler Envelope with receiver at 7m}
		\label{fig:rx_position_7m}
	\end{center}
\end{figure}

Comparing the received signal envelopes Figure \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m} the maximum Doppler increase reflects the information in Figure \ref{fig:rx_position}. The envelope in Figure \ref{fig:rx_position_0m} is also much noisier than the envelope of Figure \ref{fig:rx_position_7m} and the signal envelope, for the receiver is at 7 m, has a distinctive shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore, the receiver will be offset from the center of rotation by 7 m for the subsequent simulations.

@@

The resulting signal is analyzed by calculating its Doppler envelope. Then the maximum and minimum of that envelope will then be used to determined the overall effect of pitching the rotor blade.

%figure of doppler vs pitch
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_max_doppler.eps}
		\caption{Max and Min Envelope Frequencies vs Rotor Pitch with Tx Azimuth at 135\textdegree \space and Elevation at 54\textdegree}
		\label{fig:pitch_15_135deg}
	\end{center}
\end{figure}
%azimuth = 135deg, elevation = 54deg

@@

%figure of tx_0 and increasing pitch slope?
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_0txPos_max_doppler.eps}
		\caption{Max and Min Envelope Frequencies vs Rotor Pitch Transmitter Underneath the Helicopter}
		\label{fig:pitch_tx0}
	\end{center}
\end{figure}
%azimuth undefined, elevation 90deg

@@

%figure max doppler vs range of 15 and 7.5 pitch and 0 pitch side by side
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/elevation_angle_with_pitch_max_doppler.eps}
		\caption{Max Envelope Frequencies vs Transmitter Elevation Angle}
		\label{fig:pitch_tx_elevation_angle}
	\end{center}
\end{figure}
%azimuth = 135deg

In Figure \ref{fig:pitch_tx_elevation_angle} the maximum Doppler versus elevation angles are initially shifted up in frequency from a pitch of 0\textdegree \space and, consequently, have changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch envelopes actually become closer to the 0\textdegree \space envelope as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

%figure of 0 rx position
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Azimuth_angle_rx0_max_doppler.eps}
		\caption{Max and Min Envelope Frequencies vs Transmitter Azimuth Angle with Rx at Axis of Rotation}
		\label{fig:tx_azimuth_rx0}
	\end{center}
\end{figure}

@@

Pitching the blade modifies the results of the azimuth sweep when compared to Figure \ref{fig:tx_azimuth_angle_200}. The resulting minimum and maximum envelope frequencies versus azimuth angle for pitch are shown in Figure \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections differently according to transmitter azimuth. When the transmitter is located between 90\textdegree \space and 270\textdegree \space pitch causes an increase in Doppler but elsewhere it decreases the amount of Doppler shift. The decrease is caused by the blade surface top reflecting receding Doppler away from the receiver at those angles.

%azimuth pitch effect
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_azimuth_rev.eps}
		\caption{Max and Min Envelope Frequencies vs Transmitter Elevation Angle with 7.5deg Pitch}
		\label{fig:pitch_azimuth_rev}
	\end{center}
\end{figure}

@@

The elevation angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver, there is no vector component in the direction of rotor travel so the reflected signal is not Doppler shifted. From the model we see the amount of Doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in Figure \ref{fig:tx_elevation_135deg} where the maximum measured Doppler as the elevation angle increases.

%figure of doppler vs (elevation angle)
 \begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/elevation_angle_max_doppler_135deg.eps}
		\caption{Max and Min Envelope Frequencies vs Transmitter Elevation Angle with Tx Azimuth of 135deg}
		\label{fig:tx_elevation_135deg}
	\end{center}
\end{figure}
%azimuth = 135deg, elevation = var

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space \ref{fig:3D_model_0az_doppler}, 90\textdegree \space \ref{fig:3D_model_90az_doppler}, 180\textdegree \space \ref{fig:3D_model_180az_doppler}, and 270\textdegree \space \ref{fig:3D_model_270az_doppler} and their simulated counterparts shown in Figure \ref{fig:verify}. We see that the ray traced results match their projections on the maximum and minimum envelope Doppler as transmitter elevation angle changes. The Doppler values do not match exactly but the trends are very similar.

>>>>

After the computation of the short time Fourier transform; the Doppler profile is calculated by performing a threshold on the power, in both the positive and negative Doppler, at each time slice. The profile value for that time slice is found by approaching zero frequency and stopping as the power reaches a predetermined level from both the negative and positive values. The power level to set the threshold is calculated differently based on whether the elevation or the azimuth parameter is being swept. For the azimuth angle sweep, which will be shown later in section \ref{taa}, the maximum power is found in each time slice then those values are averaged. Then for each successive angle the same average is taken and averaged with all the others.

@@

The Doppler profiles are displayed in Figure \ref{fig:test_spec_w_doppler_profile}, the lower in blue and the upper in red. The minimum is calculated from the lower profile, and the maximum from the upper profile. Those values are then used to characterize that Doppler profile and analyze signal trends as different variables change.

@@

The rotor blade length from Figure \ref{fig:rx_position_image} is 7.5 m so the receiver will be varied from 0 m to 9 m in 1 m increments. The resulting signal is analyzed by calculating its Doppler profile. Then the maximum and minimum of that profile will then be used to determined the overall effect of the position of the receiver.

%figure of maximum doppler vs rx position
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/receiver_position_max_doppler.eps}
		\caption{Max and Min Doppler Profile vs. Receiver Position with Tx Azimuth at 135\textdegree \space and Elevation at 54\textdegree}
		\label{fig:rx_position}
	\end{center}
\end{figure}
%azimuth = 135, elevation = 54deg

Figure \ref{fig:rx_position} shows maximum Doppler versus the position of the receiver. From this graph, the maximum Doppler frequency changes significantly as the receiver is moved from directly underneath the rotor to past the tip of the rotor at 9 m.The position of the receiver also effects the resulting signal profile shape shown in Figures \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m}.

%figure of spectrogram at 0m -- figure of spectrogram at 7m side by side
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Doppler_Receiver_0m.eps}
		\caption{Doppler Profile with receiver at 0m}
		\label{fig:rx_position_0m}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Doppler_Receiver_7m.eps}
		\caption{Doppler Profile with receiver at 7m}
		\label{fig:rx_position_7m}
	\end{center}
\end{figure}

Comparing the received Doppler profiles Figure \ref{fig:rx_position_0m} and \ref{fig:rx_position_7m} the maximum Doppler increase reflects the information in Figure \ref{fig:rx_position}. The profile in Figure \ref{fig:rx_position_0m} is also much noisier than the profile of Figure \ref{fig:rx_position_7m} and the Doppler profile, for the receiver is at 7 m, has a distinctive shape with less variation between its two peaks. The position of the receiver will also determine whether the azimuth angle can be derived from the resulting signal shown later in section \ref{taa}. Therefore, the receiver will be offset from the center of rotation by 7 m for the subsequent simulations.

@@

The resulting signal is analyzed by calculating its Doppler profile. Then the maximum and minimum of that profile will then be used to determined the overall effect of pitching the rotor blade.

%figure of doppler vs pitch
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_max_doppler.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Rotor Pitch with Tx Azimuth at 135\textdegree \space and Elevation at 54\textdegree}
		\label{fig:pitch_15_135deg}
	\end{center}
\end{figure}
%azimuth = 135deg, elevation = 54deg

@@

%figure of tx_0 and increasing pitch slope?
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_0txPos_max_doppler.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Rotor Pitch Transmitter Underneath the Helicopter}
		\label{fig:pitch_tx0}
	\end{center}
\end{figure}
%azimuth undefined, elevation 90deg

@@

%figure max doppler vs range of 15 and 7.5 pitch and 0 pitch side by side
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/elevation_angle_with_pitch_max_doppler.eps}
		\caption{Max Doppler Profile Frequencies vs Transmitter Elevation Angle}
		\label{fig:pitch_tx_elevation_angle}
	\end{center}
\end{figure}
%azimuth = 135deg

In Figure \ref{fig:pitch_tx_elevation_angle} the maximum Doppler versus elevation angles are initially shifted up in frequency from a pitch of 0\textdegree \space and, consequently, have changed in slope. This is because the rays hitting the blades from below are reflecting into the receiver at an angle that is closer to the direction of blade travel. The interesting thing about this plot is that the pitch profiles actually become closer to the 0\textdegree \space profile as the elevation angle becomes smaller. Figure \ref{fig:pitch_tx_elevation_angle_difference} shows the difference plotted over elevation angle.

@@

%figure of 0 rx position
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/Azimuth_angle_rx0_max_doppler.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Transmitter Azimuth Angle with Rx at Axis of Rotation}
		\label{fig:tx_azimuth_rx0}
	\end{center}
\end{figure}

@@

Pitching the blade modifies the results of the azimuth sweep when compared to Figure \ref{fig:tx_azimuth_angle_200}. The resulting minimum and maximum profile frequencies versus azimuth angle for pitch are shown in Figure \ref{fig:pitch_azimuth_rev}. This occurs because the pitch of the blade angles RF reflections differently according to transmitter azimuth. When the transmitter is located between 90\textdegree \space and 270\textdegree \space pitch causes an increase in Doppler but elsewhere it decreases the amount of Doppler shift. The decrease is caused by the blade surface top reflecting receding Doppler away from the receiver at those angles.

%azimuth pitch effect
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/pitch_azimuth_rev.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Transmitter Elevation Angle with 7.5deg Pitch}
		\label{fig:pitch_azimuth_rev}
	\end{center}
\end{figure}

@@

The elevation angle of the transmitter in the simulation will effect the amount of Doppler added to the signal. This is due to the rays having a vector component in the direction of the travel of the rotor. When the transmitter is directly underneath the rotor and receiver, there is no vector component in the direction of rotor travel so the reflected signal is not Doppler shifted. From the model we see the amount of Doppler increase as the elevation angle decreases. This is mirrored in the simulation shown in Figure \ref{fig:tx_elevation_180deg} where the maximum measured Doppler decreases as the elevation angle increases.

%figure of doppler vs (elevation angle)
 \begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/elevation_angle_max_doppler_180deg.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Transmitter Elevation Angle with Tx Azimuth of 180deg}
		\label{fig:tx_elevation_180deg}
	\end{center}
\end{figure}

Looking back at the geometric models for transmitter azimuth angles of 0\textdegree \space \ref{fig:3D_model_0az_doppler}, 90\textdegree \space \ref{fig:3D_model_90az_doppler}, 180\textdegree \space \ref{fig:3D_model_180az_doppler}, and 270\textdegree \space \ref{fig:3D_model_270az_doppler} and their simulated counterparts shown in Figure \ref{fig:verify}. We see that the ray traced results match their projections on the maximum and minimum Doppler Profile as transmitter elevation angle changes. The Doppler values do not match exactly but the trends are very similar.
